\documentclass[a4paper,10pt]{article}
\usepackage[utf8]{inputenc}

%opening
\title{Polar lows in the ECMWF and Arome models}
\author{Matilda Hallerstig, Erik Kolstad, Linus Magnusson}

\begin{document}

\maketitle

\begin{abstract}
  Abstract
\end{abstract}

\section{Introduction}

  \subsection{Paragraph}
    Short about polar lows, what they are, why they are important, what the challenges are.
  \subsection{Paragraph}
    What the convective grey zone is, a litterature review on work on relations between convective parameterization and horizontal resolution, particularily in polar areas.
  \subsection{Paragraph}
    Short about the models (ECMWF and Arome) and why they are important. (operational in the Atlantic sector of the Arctic, increasing activity in this area)
    Also mention observations used. (scatterometer. evt. SAR)

\section{Data and methods}

\subsection{Models}
  Describe models
  \begin{itemize}
    \item table showing differences between them. (resolution, hydrostatic/non-hydrostatic, regional/global, convection parameterisations, etc)
    \item Figure with the domain of Arome Arctic 
    \item Ask Rune about Arome with deep convection?
  \end{itemize}

  \subsection{Observations}
    Description of the observation data and how it was processed
    \begin{itemize}
      \item Map with geographical coverage (Can it be combined with the map of Arome Arctic domain?)
      \item Evt. figure or something showing temporal coverage of observations?
    \end{itemize}
    
  \subsection{Diagnostic methods}
    How the tracking was done. What parameters are studied, and why.
  \subsection{Polar low cases}
    \begin{itemize}
      \item Description of polar low cases
      \item Dates and times for the model runs
      \item Maps with tracks?
    \end{itemize}

\section{Results}
  \begin{itemize} 

    \item Time series along tracks

    \item Scatterplots for an overview of model performance

    \item horizontal maps showing the structure of convective cells

    \item Vertical cross-sections to study vertical structure and dynamics of PLs?

  \end{itemize}
\section{Discussion}
  \subsection{paragraph}
    Pros and cons with different types of observations
  \subsection{paragraph}
    The role of convection and horizontal resolution, why did convection matter more in this case?
  \subsection{paragraph}
    Other differences between EC and Arome and how they might have affected the results. (global/regional, hydrostatic/nonhydrostatic)


\section{Conclusion}
  Convection mattered more than horizontal resolution.
  Explain how this knowledge is useful for further work.

\section{Acknowledgements}

\section{References}

\end{document}
